---
title: "GEMS Flux Report"
format: 
  html:
    embed-resources: true
echo: false
---

Figures for general GEMS reporting and GEMS final report.

```{r}
#| echo: false
#| message: false
#| warning: false
library(targets)
library(dplyr)
library(tidyr)
library(lubridate)
library(gemstools)
library(ggplot2)
library(plotly)
library(patchwork)
library(here)

theme_set(theme_classic())

# Configure targets to look in parent directory
tar_config_set(store = here("_targets"))

tar_load(c(
  rga_binned,
  rga_binned_wide,
  hourly_flux,
  hourly_stats,
  monthly_stats,
  monthly_nem,
  daily_nem,
  rga_oxygen,
  seaphox_df_jul,
  proco2_df_jul,
  ox_model,
  ox_cal_df,
  eelgrass,
  par_model_df
))
```

# RGA timeseries


O2/Ar representative plot

```{r}
oxp <- rga_binned |>
  filter(
    timestamp > as.POSIXct("2025-07-15 00:00:00"),
    timestamp < as.POSIXct("2025-07-20 00:00:00")
  ) |>
  ggplot(aes(timestamp, mass_32_40, color = inlet)) +
  geom_line() +
  labs(
    title = "Oxygen/Argon ratio at high and low inlets",
    x = NULL,
    y = "Mass 32:40"
  )

oxp
```

Oxygen/Argon timeseries

```{r}
rga_binned_wide |>
  arrange(timestamp) |>
  select(timestamp, mass_32_40_high, mass_32_40_low) |>
  mutate(
    time_diff = as.numeric(difftime(
      timestamp,
      lag(timestamp),
      units = "hours"
    )),
    mass_32_40_high = ifelse(time_diff > 1, NA, mass_32_40_high),
    mass_32_40_low = ifelse(time_diff > 1, NA, mass_32_40_low)
  ) |>
  pivot_longer(
    cols = c(mass_32_40_low, mass_32_40_high),
    names_to = "inlet",
    names_prefix = "mass_32_40_",
    values_to = "mass_32_40"
  ) |>
  ggplot(aes(timestamp, mass_32_40, color = inlet)) +
  geom_line(na.rm = TRUE)
```

## RGA 

### Mass traces

RGA signals for masses of interest.
Each mass is measured approximately every 10 s.
A five day period of the deployment was chosen
to highlight diel cycles. Each signal is normalized
to mass 40 (Argon) to control for effects
of temperature, pressure, and biofouling.
The mass 34 signal is likely an interference
with Oxygen, rather than a true hydrogen sufide signal.

TODO: mass units
TODO: use 15 min averages?
  
```{r}
#| fig-height: 8
jul_rga <- rga_binned |>
  filter(
    timestamp > as.POSIXct("2025-07-15 00:00:00"),
    timestamp < as.POSIXct("2025-07-20 00:00:00")
  )

mep <- jul_rga |>
  ggplot(aes(timestamp, mass_15_40, color = inlet)) +
  geom_line() +
  labs(
    title = "Methane (mass 15)",
    x = NULL,
    y = "Mass 15:40"
  )

nip <- jul_rga |>
  ggplot(aes(timestamp, mass_28_40, color = inlet)) +
  geom_line() +
  labs(
    title = "Nitrogen (mass 28)",
    x = NULL,
    y = "Mass 28:40"
  ) +
  theme(legend.position = "none")

sup <- jul_rga |>
  ggplot(aes(timestamp, mass_34_40, color = inlet)) +
  geom_line() +
  labs(
    title = "Hydrogen Sulfide (mass 34)",
    x = NULL,
    y = "Mass 34:40"
  ) +
  theme(legend.position = "none")


cop <- jul_rga |>
  ggplot(aes(timestamp, mass_44_40, color = inlet)) +
  geom_line() +
  labs(
    title = "Carbon Dioxide (mass 44)",
    x = NULL,
    y = "Mass 44:40"
  ) +
  theme(legend.position = "none")


oxp <- oxp +
  ggtitle("Oxygen (mass 32)") +
  theme(legend.position = "none")


mep / nip / oxp / sup / cop
```

### Effect of Argon normalization

Normalizing measured masses to Argon (mass 40)
controls for physical effects on gas solubility and membrane transport.
Gas transport is affected by temperature, pressure and water flow rate.
This figure shows normalized and un-normalized Nitrogen (mass 28)
across a period of increasing biofouling at the high inlet,
followed by cleaning the inlets on Sept. 11.

```{r}
sep_rga <- rga_binned |>
  filter(
    timestamp > as.POSIXct("2025-09-04 00:00:00"),
    timestamp < as.POSIXct("2025-09-14 00:00:00")
  )

nup <- sep_rga |>
  ggplot(aes(timestamp, mass_28, color = inlet)) +
  geom_line() +
  labs(
    title = "Nitrogen (mass 28, unnormalized)",
    x = NULL,
    y = "Mass 28"
  )


nnp <- sep_rga |>
  ggplot(aes(timestamp, mass_28_40, color = inlet)) +
  geom_line() +
  labs(
    title = "Nitrogen (normalized to Argon)",
    x = NULL,
    y = "Mass 28:40"
  ) +
  theme(legend.position = "none")

nup / nnp
```

# Calibration

TODO: calibration in umol/l

```{r}
om <- ox_model$coefficients[2]
oi <- ox_model$coefficients[1]
r2 <- summary(ox_model)$r.squared

cal_plot_df <- ox_cal_df |>
  filter(timestamp > "2025-07-11 10:00:00") |>
  mutate(
    rga_ox = mass_32_40 * om + oi
  )
```

### SeapHOx vs calibrated GEMS Oxygen

```{r}
#| warning: false

eq_label <- paste0(
  "y = ",
  round(oi, 2),
  " + ",
  round(om, 2),
  "x\n",
  "R² = ",
  round(r2, 3)
)

ggplot(cal_plot_df, aes(mass_32_40, seaphox_oxygen_ml_l)) +
  geom_smooth(method = "lm", se = FALSE, color = "darkgrey") +
  annotate(
    "text",
    x = Inf,
    y = -Inf,
    label = eq_label,
    hjust = 1.1,
    vjust = -1
  ) +
  geom_point(
    shape = 1,
    color = "blue",
    size = 2,
    stroke = 0.5,
    alpha = 0.2
  ) +
  #xlim(3, 9) +
  #ylim(3, 9) +
  labs(
    #title = "Calibration fit",
    y = "Optode oxygen [ml/l]",
    x = "GEMS mass 32:40"
  )
```

### Fitted Oxygen

```{r}
cal_plot_df |>
  pivot_longer(
    cols = c(seaphox_oxygen_ml_l, rga_ox),
    names_to = "source",
    values_to = "oxygen_ml_l"
  ) |>
  mutate(source = ifelse(source == "rga_ox", "GEMS", "Optode")) |>
  ggplot(aes(timestamp, oxygen_ml_l, color = source)) +
  geom_line() +
  labs(title = "Calibration timeseries", x = NULL, y = "Oxygen [ml/l]")
```

# Gradient

Gradient plots

# Flux
  
## Hourly oxygen flux timeseries

Not showing CO2 flux for now.

```{r}
gop <- hourly_flux |>
  mutate(
    time_diff = as.numeric(difftime(
      timestamp,
      lag(timestamp),
      units = "hours"
    )),
    ox_flux = ifelse(time_diff > 1, NA, ox_flux),
  ) |>
  ggplot(aes(timestamp, ox_flux)) +
  geom_hline(yintercept = 0) +
  geom_line(na.rm = TRUE) +
  labs(title = "Hourly Oxygen Flux", x = NULL, y = "Oxygen flux [mmol/m²/h]")

gcp = ggplot(hourly_flux, aes(timestamp, co2_flux)) +
  geom_line()

tp <- hourly_flux |>
  mutate(
    time_diff = as.numeric(difftime(
      timestamp,
      lag(timestamp),
      units = "hours"
    )),
    adv_temp_mean = ifelse(time_diff > 1, NA, adv_temp_mean),
  ) |>
  ggplot(aes(timestamp, adv_temp_mean)) +
  geom_line() +
  labs(title = "Temperature", x = NULL, y = "Temp [C]")

gpp = ggplot(hourly_flux, aes(timestamp, par_mean)) +
  geom_line() +
  labs(title = "Predicted PAR", x = NULL, y = "Predicted PAR [µmol/m²/s¹]")

# DLI daily PAR
min_flux_ts <- as.Date(min(hourly_flux[["timestamp"]]))
max_flux_ts <- as.Date(max(hourly_flux[["timestamp"]]))

par_model_df <- par_model_df |>
  filter(date >= min_flux_ts, date <= max_flux_ts)

dpp <- ggplot(par_model_df, aes(date, dli_mol_m2_day)) +
  geom_line() +
  labs(
    title = "Predicted Daily PAR",
    x = NULL,
    y = "Predicted DLI [mol/m²/day¹]"
  )

gop / dpp / tp
```

Representative daily flux? July 16-19?

```{r}
jul_flux <- hourly_flux |>
  filter(
    timestamp > as.POSIXct("2025-07-15 00:00:00"),
    timestamp < as.POSIXct("2025-07-19 12:00:00")
  )
jfp <- jul_flux |>
  ggplot(aes(timestamp, ox_flux)) +
  geom_hline(yintercept = 0) +
  geom_line() +
  labs(
    title = "July Hourly Oxygen Flux",
    x = NULL,
    y = "Oxygen flux [mmol/m²/h]"
  )

jpp <- jul_flux |>
  ggplot(aes(timestamp, par_mean)) +
  geom_line() +
  labs(title = "Predicted PAR", x = NULL, y = "Predicted PAR [µmol/m²/s¹]")

jfp / jpp
```

## Diel Plots

```{r}
hmp <- hourly_stats |>
  ggplot(aes(solar_hour, ox_mean_umol_l_mean_mean)) +
  geom_line() +
  geom_pointrange(
    aes(
      ymin = ox_mean_umol_l_mean_mean - ox_mean_umol_l_mean_se,
      ymax = ox_mean_umol_l_mean_mean + ox_mean_umol_l_mean_se
    ),
    size = 0.1,
    # position = position_dodge(width = 0.5)
  ) +
  labs(x = NULL, y = "Oxygen Conc. [mmol/m3]")

hgp <- hourly_stats |>
  ggplot(aes(solar_hour, ox_gradient_umol_l_m_mean_mean)) +
  geom_line() +
  geom_pointrange(
    aes(
      ymin = ox_gradient_umol_l_m_mean_mean - ox_gradient_umol_l_m_mean_se,
      ymax = ox_gradient_umol_l_m_mean_mean + ox_gradient_umol_l_m_mean_se
    ),
    size = 0.1,
    # position = position_dodge(width = 0.5)
  ) +
  geom_hline(yintercept = 0) +
  labs(x = "Hour of Day", y = "Oxygen Gradient [mmol/m4]")

flp <- hourly_stats |>
  ggplot(aes(solar_hour, ox_flux_mean)) +
  geom_line() +
  geom_pointrange(
    aes(
      ymin = ox_flux_mean - ox_flux_se,
      ymax = ox_flux_mean + ox_flux_se
    ),
    size = 0.1,
    # position = position_dodge(width = 0.5)
  ) +
  geom_hline(yintercept = 0) +
  labs(x = NULL, y = "Oxygen Flux [mmol/m2/h]")

parp <- hourly_stats |>
  ggplot(aes(solar_hour, par_mean_mean)) +
  geom_line() +
  labs(x = "Hour of Day", y = "Predicted PAR [mmol/m2]")
```

### Concentration, gradient

```{r}
hmp / hgp

# hm_ply <- ggplotly(hmp, dynamicTicks = TRUE)
# hg_ply <- ggplotly(hgp, dynamicTicks = TRUE)
# subplot(hm_ply, hg_ply, nrows = 2, shareX = TRUE, titleY = TRUE)
```

### Oxygen flux and predicted par

```{r}
flp / parp
```

### Monthly variability

Same as above but plotted by month

June removed due to incomplete data

Oct removed for now due to inlet issue

```{r}

monthly_stats <- mutate(monthly_stats, month = as.factor(month))

hmp <- monthly_stats |>
  ggplot(aes(solar_hour, ox_mean_umol_l_mean_mean, color = month)) +
  geom_line() +
  geom_pointrange(
    aes(
      ymin = ox_mean_umol_l_mean_mean - ox_mean_umol_l_mean_se,
      ymax = ox_mean_umol_l_mean_mean + ox_mean_umol_l_mean_se
    ),
    size = 0.1,
    # position = position_dodge(width = 0.5)
  ) +
  geom_hline(yintercept = 0) +
  labs(x = NULL, y = "Oxygen Conc. [mmol/m3]")

hgp <- monthly_stats |>
  ggplot(aes(solar_hour, ox_gradient_umol_l_m_mean_mean, color = month)) +
  geom_line() +
  geom_pointrange(
    aes(
      ymin = ox_gradient_umol_l_m_mean_mean - ox_gradient_umol_l_m_mean_se,
      ymax = ox_gradient_umol_l_m_mean_mean + ox_gradient_umol_l_m_mean_se
    ),
    size = 0.1,
    # position = position_dodge(width = 0.5)
  ) +
  geom_hline(yintercept = 0) +
  labs(x = "Hour of Day", y = "Oxygen Gradient [mmol/m4]")

flp <- monthly_stats |>
  ggplot(aes(solar_hour, ox_flux_mean, color = month)) +
  geom_line() +
  geom_pointrange(
    aes(
      ymin = ox_flux_mean - ox_flux_se,
      ymax = ox_flux_mean + ox_flux_se
    ),
    size = 0.1,
    # position = position_dodge(width = 0.5)
  ) +
  geom_hline(yintercept = 0) +
  labs(x = NULL, y = "Oxygen Flux [mmol/m2/h]")

parp <- monthly_stats |>
  ggplot(aes(solar_hour, par_mean_mean, color = month)) +
  geom_line() +
  labs(x = "Hour of Day", y = "Predicted PAR [mmol/m2]")

flp / parp

# hm_ply <- ggplotly(hmp, dynamicTicks = TRUE)
# hg_ply <- ggplotly(hgp, dynamicTicks = TRUE)
# subplot(hm_ply, hg_ply, nrows = 2, shareX = TRUE, titleY = TRUE)
```


# Flux barplots

monthly NEM of diel curves
Include light, temp

```{r}
ggplot(monthly_nem, aes(month, nem_mmol_m2_day)) +
  geom_col() +
  geom_hline(yintercept = 0) +
  labs(title = "Monthly Net Ecosystem Metabolism", y = "NEM [mmol m2 day]")
```

```{r}
ggplot(daily_nem, aes(day, nem_mmol_m2_h)) +
  geom_col() +
  geom_hline(yintercept = 0) +
  labs(title = "Daily Net Ecosystem Metabolism", y = "NEM [mmol m2 h]")
```

# Eelgrass growth figures

Plots below are Naushon eelgrass quadrat data plotted by year. Point error bars are SD of quadrats for mass and count, SD of all blade lengths for length. Fit line is a loess fit through all data.

```{r}
bmp <- ggplot(eelgrass, aes(day_of_year, density_g_m2, color = year)) +
  geom_smooth(inherit.aes = FALSE, aes(day_of_year, density_g_m2), se = FALSE) +
  geom_line() +
  geom_pointrange(aes(
    ymin = density_g_m2 - density_g_m2_sd,
    ymax = density_g_m2 + density_g_m2_sd
  )) +
  scale_x_continuous(
    breaks = c(1, 32, 60, 91, 121, 152, 182, 213, 244, 274, 305, 335),
    labels = month.abb
  ) +
  labs(title = "Eelgrass Shoot Biomass", x = NULL, y = "Density [g/m2]")

blp <- ggplot(eelgrass, aes(day_of_year, length_cm, color = year)) +
  geom_smooth(inherit.aes = FALSE, aes(day_of_year, length_cm), se = FALSE) +
  geom_line() +
  geom_line() +
  geom_pointrange(aes(
    ymin = length_cm - length_cm_sd,
    ymax = length_cm + length_cm_sd
  )) +
  scale_x_continuous(
    breaks = c(1, 32, 60, 91, 121, 152, 182, 213, 244, 274, 305, 335),
    labels = month.abb
  ) +
  labs(title = "Eelgrass Shoot length", x = NULL, y = "Length [cm]")

bmp / blp
```


## mass

## count

## length

# Synthesis figures

flux vs light, growth, temp
